Converging High-Throughput and High-Performance Computing: A Case Study

### 2017-03-28 overall.tex
Iteration PanDA I/O Impact at OLCF

To better understand the I/O impact of ATLAS PanDA project on Titan
supercomputing environment we analyzed 1,175 jobs ran on the week of 10/25/2016,
for a total of 174 hours. Table~\ref{panda-olcf-stats} shows the overall
statistical breakdown of the observed file I/O impact of ATLAS at OLCF\@.
Figures~\ref{fig:atlas-titan-io-read} and~\ref{fig:atlas-titan-io-written} show
the file read and write I/O histograms for these 1,175 jobs, respectively.
Figures~\ref{fig:atlas-titan-file-open} and~\ref{fig:atlas-titan-file-close}
show the file $open()$ and $close()$ metadata load histograms of the same 1,175
ATLAS jobs, respectively.

As can be seen from Table~\ref{panda-olcf-stats}, the number of nodes used by
ATLAS jobs vary between 1 and 300, while the average is at 35. 75\% of the ATLAS
jobs consume less than 25 and 92\% consume less than or equal to 100 Titan
compute nodes. During the 174 hours of data collection, we observed that 6.75
ATLAS jobs were executed on average per hour on Titan and ran for 1.74 hours on
average.

ATLAS jobs issues a large number of file read operations, as can be seen in
Table~\ref{panda-olcf-stats}. The maximum amount read by any ATLAS job in
aggregate in this observed period was less than 250 GB and the maximum amount
written in aggregate was less than 75 GB\@. The average amount read per job is
20 GB and average amount written is 6 GB\@.

\begin{figure}[!htb]
    \centering
    \begin{tabular}{cc}
        {\includegraphics[width=0.48\textwidth]{figures/panda_data_read_finer_hist.pdf}}\\
    \end{tabular}
    \caption{ATLAS file read operation histogram on Titan for week of 10/25/16.}
\label{fig:atlas-titan-io-read}
\end{figure}

\begin{figure}[!htb]
    \centering
    \begin{tabular}{cc}
        {\includegraphics[width=0.48\textwidth]{figures/panda_data_written_finer_hist.pdf}}\\
    \end{tabular}
    \caption{ATLAS file write I/O histogram on Titan for week of 10/25/16.}
\label{fig:atlas-titan-io-written}
\end{figure}

Per job read and write file I/O statistics show an interesting pattern.
Table~\ref{panda-olcf-stats} indicates that the amount of data read per ATLAS
compute node on Titan is less than 400 MB on average, while the amount of data
written per node is less than 170 MB on average. This correlates with our
finding that ATLAS PanDA jobs are read heavy. However, as can be seen in
Table~\ref{panda-olcf-stats} and figures {fig:atlas-titan-io-read} and
{fig:atlas-titan-io-written}, the distribution between read and written amount
of data per job is quite different from one another. The read operation
distribution per job shows a long tail, ranging from 12.5 GB to 250 GB, while
the written amount of data has a very narrow distribution.

\begin{figure}[!htb]
    \centering
    \begin{tabular}{cc}
        {\includegraphics[width=0.48\textwidth]{figures/panda_file_open_hist.pdf}}\\
    \end{tabular}
    \caption{ATLAS file $open()$ histogram on Titan for week of 10/25/16.}
\label{fig:atlas-titan-file-open}
\end{figure}

\begin{figure}[!htb]
    \centering
    \begin{tabular}{cc}
        {\includegraphics[width=0.48\textwidth]{figures/panda_file_close_hist.pdf}}\\
    \end{tabular}
    \caption{ATLAS file $close()$ histogram on Titan for week of 10/25/16.}
\label{fig:atlas-titan-file-close}
\end{figure}

On the metadata I/O breakdown, ATLAS PanDA jobs yield 23 file $open()$
operations and 5 file $close()$ operations per second. The file $open()$
operations listed here don't include the file $stat()$ operations. As can be
seen from figures~\ref{fig:atlas-titan-file-open}
and~\ref{fig:atlas-titan-file-close}, they exhibit similar distributions. The
maximum number of file $open()$ operations are around 170 on average and the
maximum number of file $close()$ operations are 39 on average per second per
job. The total number of file $open()$ operations is 172,089,760 for the
observed window of 174 hours, while the total number of file $close()$
operations stand at 40,132,992 for the total 1,175 ATLAS PanDA jobs in the same
observation window. The difference between these two values is puzzling and it
is under investigation at the time of writing this paper. One possible
explanation is that ATLAS PanDA jobs perhaps don't call a file $close()$
operation per every file $open()$ issued.

@@

Overall, based on our experiments with real-world jobs, it can be safely
concluded that the file and metadata I/O load of ATLAS PanDA project on the OLCF
Titan supercomputing environment and the Spider 2 file system is not detrimental
to the center operations and the overall impact is minimal, at the current scale
of the project.

>>>>

To better understand the I/O impact of ATLAS PanDA project on Titan
supercomputing environment we analyzed 1,175 jobs ran on the week of 10/25/2016,
for a total of 174 hours. Table~\ref{panda-olcf-stats} shows the overall
statistical breakdown of the observed file I/O impact of ATLAS at OLCF\@.
% Figures~\ref{fig:atlas-titan-io-read} and~\ref{fig:atlas-titan-io-written}
% show the file read and write I/O histograms for these 1,175 jobs,
% respectively. Figures~\ref{fig:atlas-titan-file-open}
% and~\ref{fig:atlas-titan-file-close} show the file $open()$ and $close()$
% metadata load histograms of the same  1,175 ATLAS jobs, respectively.
%
% As can be seen from Table~\ref{panda-olcf-stats},

ATLAS jobs use between 1 and 300 worker nodes, and 35 on average. 75\% of the
ATLAS jobs consume less than 25 nodes and 92\% less than 100. During the 174
hours of data collection, 6.75 ATLAS jobs were executed on average per hour,
each job running for an average of 1.74 hours.
% ATLAS jobs issues a large number of file read operations, as can be seen in
% Table~\ref{panda-olcf-stats}.
Every job read less than 250 GB and wrote less than 75 GB of data and, on
average, each job read 20 GB and wrote 6 GB of data.

@@

% \begin{figure}[!htb]
%     \centering
%     \begin{tabular}{cc}
%         {\includegraphics[width=0.48\textwidth]{figures/panda_data_read_finer_hist.pdf}}\\
%     \end{tabular}
%     \caption{ATLAS file read operation histogram on Titan for week of 10/25/16.}
% \label{fig:atlas-titan-io-read}
% \end{figure}
%
% \begin{figure}[!htb]
%     \centering
%     \begin{tabular}{cc}
%         {\includegraphics[width=0.48\textwidth]{figures/panda_data_written_finer_hist.pdf}}\\
%     \end{tabular}
%     \caption{ATLAS file write I/O histogram on Titan for week of 10/25/16.}
% \label{fig:atlas-titan-io-written}
% \end{figure}

The I/O of the ATLAS job executed on Titan show an interesting pattern.
Table~\ref{panda-olcf-stats} indicates that the amount of data read per ATLAS
worker node is less than 400 MB on average, while the amount of data written per
node is less than 170 MB on average. This correlates with our finding that ATLAS
PanDA jobs are read heavy. However,

ATLAS PanDA jobs are read heavy: On average, the amount of data read per worker
node is less than 400 MB, while the amount of data written is less than 170 MB.
This correlates with our finding that . However,
% as can be seen in Table~\ref{panda-olcf-stats} and figures
% {fig:atlas-titan-io-read} and {fig:atlas-titan-io-written},
the distributions of read and written data are quite different: The read
operation distribution per job shows a long tail, ranging from 12.5 GB to 250
GB, while the written amount of data has a very narrow distribution.\mtnote{Should we say why?}

% \begin{figure}[!htb]
%     \centering
%     \begin{tabular}{cc}
%         {\includegraphics[width=0.48\textwidth]{figures/panda_file_open_hist.pdf}}\\
%     \end{tabular}
%     \caption{ATLAS file $open()$ histogram on Titan for week of 10/25/16.}
% \label{fig:atlas-titan-file-open}
% \end{figure}
%
% \begin{figure}[!htb]
%     \centering
%     \begin{tabular}{cc}
%         {\includegraphics[width=0.48\textwidth]{figures/panda_file_close_hist.pdf}}\\
%     \end{tabular}
%     \caption{ATLAS file $close()$ histogram on Titan for week of 10/25/16.}
% \label{fig:atlas-titan-file-close}
% \end{figure}

The metadata I/O breakdown shows that ATLAS jobs yield 23 file $open()$
operations per second (not including file $stat()$ operations) and 5 file $close()$ operations per second, with similar distributions.
% As can be seen from figures~\ref{fig:atlas-titan-file-open}
% and~\ref{fig:atlas-titan-file-close}, they .
On average, the maximum number of file $open()$ operations per job is
$\approx$170/s and the maximum number of file $close()$ operations is
$\approx$39/s.
% on average per second per job.
For the 1,175 ATLAS PanDA jobs observed, the total number of file $open()$
operations is 172,089,760 and the total number of file $close()$ operations is
40,132,992. The difference between these two values is still under
investigation: One possible explanation is that ATLAS PanDA jobs don't call a
file $close()$ operation per every file $open()$ issued.

Overall, based on our experiments with ATLAS jobs, it can be safely concluded
that the file and metadata I/O load of ATLAS PanDA project on the OLCF Titan
supercomputing environment and the Spider 2 file system is not detrimental to
the center operations. At the current scale of the project, the overall impact
of ATLAS operations on Titan is minimal.
